### 2021-03-19 analysis/01_clean-data.Rmd
Clean and process data

---
title: "Process and merge data"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knit_print.data.frame <- function(x, ...) {
  res <- paste(c('', '', kable_styling(kable(x, booktabs = TRUE))), collapse = '\n')
  asis_output(res)
}

registerS3method("knit_print", "data.frame", knit_print.data.frame)
registerS3method("knit_print", "grouped_df", knit_print.data.frame)

knitr::opts_chunk$set(fig.retina = 3,
                      tidy.opts = list(width.cutoff = 120),  # For code
                      options(width = 90),  # For output
                      fig.asp = 0.618, fig.width = 7, 
                      fig.align = "center", out.width = "85%")

options(dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(countrycode)
library(here)
```

# Load raw data

```{r load-raw-data, cache=TRUE, warning=FALSE, message=FALSE}
# V-Dem pandemic backsliding
pandem_raw <- read_csv(here("data", "raw_data", "pandem", "csv_files", "PanDem_ts_V5.csv"))

# ICCPR derogations
derogations_raw <- read_csv(here("data", "raw_data", "derogations", "Pandemic backsliding - Sheet1.csv"))

# V-Dem
vdem_raw <- read_rds(here("data", "raw_data", "Country_Year_V-Dem_Full+others_R_v10",
                         "V-Dem-CY-Full+Others-v10.rds")) %>% as_tibble()
```

# Clean up data

```{r clean-data}
emlaw_lookup <- tribble(
  ~emlaw, ~emlaw_fct, ~emlaw_cond,
  1, "Declaration in existing legal framework", "Emergency response with legal instruments",
  2, "Declaration in existing legal framework, distinguishes public health emergency", "Emergency response with legal instruments",
  3, "Declaration of disaster where different from emergency", "Emergency response with legal instruments",
  4, "Declaration using other legislation", "Emergency response with legal instruments",
  5, "Emergency response without legal instruments", "Emergency response without legal instruments",
  6, "No emergency response", "No emergency response",
  7, "Other", "Other"
)

regime_lookup <- tribble(
  ~v2x_regime, ~regime,
  0, "Closed autocracy",
  1, "Electoral autocracy",
  2, "Electoral democracy",
  3, "Liberal democracy"
) %>% 
  mutate(regime = fct_inorder(regime, ordered = TRUE))

type_lookup <- tribble(
  ~type, ~type_clean,
  "A", "Derogated as intended",
  "B", "Derogated; measures not temporary, proportional, or necessary",
  "C", "Emergency declared; no formal derogation",
  "D", "No emergency; no formal derogation",
  "E", "Has not ratified ICCPR"
) %>% 
  mutate(type_clean_letter = paste0(type, ": ", type_clean)) %>% 
  mutate(across(starts_with("type"), ~fct_inorder(., ordered = TRUE)))

pandem_clean <- pandem_raw %>% 
  mutate(ndrights_fct = factor(ndrights, levels = 0:1, 
                               labels = c("No violations", "Violations of non-derogable rights"))) %>% 
  group_by(country_name) %>% 
  mutate(emlaw_min = min(emlaw),
         pandem_max = max(pandem),
         panback_max = max(panback)) %>% 
  ungroup() %>% 
  left_join(emlaw_lookup, by = "emlaw") %>% 
  left_join(rename_with(emlaw_lookup, ~str_c(., "_min"), everything()), by = "emlaw_min") %>% 
  left_join(regime_lookup, by = "v2x_regime") %>% 
  mutate(emlaw_fct = factor(emlaw_fct, levels = emlaw_lookup$emlaw_fct, ordered = TRUE),
         emlaw_cond = factor(emlaw_cond, levels = unique(emlaw_lookup$emlaw_cond), ordered = TRUE)) %>% 
  mutate(emlaw_min_fct = factor(emlaw_fct, levels = emlaw_lookup$emlaw_fct, ordered = TRUE),
         emlaw_min_cond = factor(emlaw_cond, levels = unique(emlaw_lookup$emlaw_cond), ordered = TRUE)) %>% 
  mutate(cowcode = countrycode(country_name, "country.name", "cown",
                               custom_match = c("Serbia" = 345,
                                                "Hong Kong" = 997)),
         iso3 = countrycode(country_name, "country.name", "iso3c"))

pandem_single <- pandem_clean %>% 
  filter(Quarter == "Q4")
glimpse(pandem_single)

derogations_clean <- derogations_raw %>% 
  rename(ratified = `Did they ratify/are they a member? (1=Yes, 0=No, 99= Not possible)`,
         derogate = `Did they officially derogate?`,
         derogation_start = `Derogation start date`,
         derogation_end = `Derogation end date`,
         country_name = Country) %>% 
  mutate(country_name = recode(country_name, "Columbia" = "Colombia")) %>% 
  mutate(cowcode = countrycode(country_name, "country.name", "cown",
                               custom_match = c("Serbia" = 345,
                                                "State of Palestine" = 998,
                                                "Cook Islands" = NA,
                                                "Niue" = NA)),
         iso3 = countrycode(country_name, "country.name", "iso3c"))
glimpse(derogations_clean)

vdem_clean <- vdem_raw %>% 
  filter(year > 2015) %>% 
  rename(cowcode = COWcode) %>% 
  mutate(cowcode = case_when(
    country_name == "Hong Kong" ~ 997,
    country_name == "Palestine/West Bank" ~ 998,
    TRUE ~ cowcode
  )) %>% 
  mutate(iso3c = countrycode(country_name, "country.name", "iso3c",
                             custom_match = c("Kosovo" = "XKK",
                                              "Somaliland" = "XSO",
                                              "Zanzibar" = "EAZ"))) %>% 
  select(country_name, year, cowcode, iso3c, 
         # Civil society stuff
         v2cseeorgs,  # CSO entry and exit
         v2csreprss,  # CSO repression
         v2cscnsult,  # CSO consultation
         v2csprtcpt,  # CSO participatory environment
         v2csgender,  # CSO women's participation
         v2csantimv,  # CSO anti-system movements
         v2xcs_ccsi,  # Core civil society index (entry/exit, repression, participatory env)
         # Human rights and politics
         # Political corruption index (less to more, 0-1) (public sector +
         # executive + legislative + judicial corruption)
         v2x_corr,
         # Rule of law index
         v2x_rule,
         # Rights indexes
         v2x_civlib,  # Civil liberties index
         v2x_clphy,  # Physical violence index
         v2x_clpriv,  # Private civil liberties index
         v2x_clpol,  # Political civil liberties index
         # Democracy
         v2x_polyarchy
  )
glimpse(vdem_clean)

pandem_derog <- derogations_clean %>% 
  left_join(pandem_single, by = c("country_name", "cowcode", "iso3")) %>% 
  left_join(filter(vdem_clean, year == 2019), by = c("country_name", "cowcode")) %>% 
  filter(!is.na(X1)) %>%
  mutate(has_derogation_end = case_when(
    is.na(derogation_end) ~ FALSE,
    str_detect(derogation_end, "N/A") ~ FALSE,
    TRUE ~ TRUE
  )) %>% 
  mutate(has_emergency_end = emlimit == 1) %>% 
  mutate(nonproportional = Type1 > 1 | Type2 == 1 | Type3 > 1) %>% 
  mutate(type = case_when(
    derogate == 1 & has_derogation_end & !nonproportional ~ "A",
    derogate == 1 & (!has_derogation_end | nonproportional) ~ "B",
    derogate == 0 & ratified == 1 & emlaw < 5 ~ "C",
    derogate == 0 & ratified == 1 & emlaw >= 5 ~ "D",
    TRUE ~ "E"
  )) %>% 
  left_join(type_lookup, by = "type")
glimpse(pandem_derog)
```

# Save data for later use

```{r save-data}
saveRDS(pandem_clean, here("data", "derived_data", "pandem_clean.rds"))
saveRDS(pandem_single, here("data", "derived_data", "pandem_single.rds"))
saveRDS(derogations_clean, here("data", "derived_data", "derogations_clean.rds"))
saveRDS(vdem_clean, here("data", "derived_data", "vdem_clean.rds"))
saveRDS(pandem_derog, here("data", "derived_data", "pandem_derog.rds"))
```